---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loaddata}
library(data.table)
data <- read.csv("activity.csv", header = TRUE)
data$date <- as.Date(strptime(data$date, "%Y-%m-%d"))
head(data)
```

## What is mean total number of steps taken per day?
```{r mean}
stepsperday <- aggregate(.~date,data=data[,1:2],sum)
head(stepsperday)
barplot(stepsperday$steps, las=2, names.arg=stepsperday$date)
title(main="Total steps per each day")
mean(stepsperday$steps)
median(stepsperday$steps)
```

## What is the average daily activity pattern?
```{r pattern}
avgstepsperint <- aggregate(.~interval, data=data[,c(1,3)], mean)
head(avgstepsperint)
plot(avgstepsperint$interval, avgstepsperint$steps, type="l", ylab="", xlab="", las=1)
title(main="Average daily activity pattern", ylab="Average steps", xlab="Intervals")
```

#### The following 5-minute interval contains the maximum number of steps:
```{r maxsteps}
avgstepsperint$interval[avgstepsperint$steps == max(avgstepsperint$steps)]
```

## Imputing missing values

#### Total number of missing values in the data is 2304 as you can see in the below result:
```{r numNA}
missing <- is.na(data)
table(missing)
```

#### The following few steps will create a new dataset which is the same as the original, but all the NAs are replaced by the mean of the steps per each interval:
```{r fillinginmissing}
tempdata <- merge(subset(data, missing), avgstepsperint, by = "interval")
tempdata <- tempdata[order(tempdata[,3], tempdata[,1]),]
head(tempdata)
data2 <- data
head(data2)
data2[missing] <- tempdata[,4]
missing2 <- is.na(data2)
table(missing2)
head(data2)
```

#### Total number of steps per day taken based on the new dataset
```{r mean2}
stepsperday2 <- aggregate(.~date,data=data2[,1:2],sum)
head(stepsperday2)
barplot(stepsperday2$steps, las=2, names.arg=stepsperday2$date)
title(main="Total steps per each day")
mean(stepsperday2$steps)
median(stepsperday2$steps)
```

#### What is the impact of imputing missing data on the estimates of the total daily number of steps?
Total daily number of steps have been increased on those days where any NAs were replaced by the mean of the average steps per the particular interval.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
Sys.setlocale("LC_TIME", "C")
day <- factor(c("weekday", "weekend"))
data2[weekdays(data2$date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),4] <- "weekday"
data2[weekdays(data2$date) %in% c("Sunday", "Saturday"),4] <- "weekend"
head(data2)
table(data2$day)
```